---
title: "Market Overview"
author: "Vi Duong"
date: "2024-03-24"
format:
  html:
    embed-resources: false
    toc: true
    number-sections: true
    theme: Yeti
    fontsize: 1.1em
    linestretch: 1.7
---

```{r}
#| echo: false
#| include: false

library(pacman)
p_load(dplyr
       , readxl
       , janitor
       , lubridate
       , stringr
       , DT
       , purrr
       , highcharter
       , tidyr
       , reactable
       , tibble
       , networkD3
       , RDataOps
       , sf
       , knitr
       , htmltools
       , leaflet
       , plotly
       , googlesheets4
       , kableExtra
       , formattable
       , cohorts
       , openxlsx
       , RColorBrewer)

source("../source_functions.R")

# Color
pink <- "#D81159"
lightpink <- "#FF6893"
lightgreen <- "#60BAAF"
green <- "#00524A"


```

```{r, include=FALSE}
load("../../../../[02] Ads Optimization/DATA/[02] Clean Data/gmv_nmv.Rdata")
load("../../../../[02] Ads Optimization/DATA/[02] Clean Data/competitor_data_2023.Rdata")
load("../../../../[02] Ads Optimization/DATA/[02] Clean Data/competitor_data_2022_2023.Rdata")
load("../../../../[02] Ads Optimization/DATA/[02] Clean Data/competitor_shopee_cate.Rdata")
load("../../../../[02] Ads Optimization/DATA/[02] Clean Data/competitor_tiktok_cate.Rdata")

tiktok_rev <- ReadExcelFiles("../../../../[02] Ads Optimization/DATA/[00] Raw Data/tiktok_last6months_gmv.xlsx")

link <- "https://docs.google.com/spreadsheets/d/1uuw0D6UYpjpZMLAOPfkzJM-8MoP4yKujC0iaE1rQkRE/edit#gid=0"

```

```{r, include=FALSE}
gmv_nmv <- gmv_nmv %>%
  mutate(est_gmv = case_when(gmv <= nmv ~ nmv * (1+0.53)
                             , TRUE ~ gmv)) %>% 
  filter(est_gmv > 0) %>% 
  mutate(creator_type = ifelse(is.na(creator_type), "partner", creator_type)) %>% 
  mutate(category = case_when(category %in% c("Womenswear", "womenswear") ~ "Trang phục nữ & Đồ lót"
                              , category %in% c("Menswear") ~ "Trang phục nam & Đồ lót"
                              , category %in% c("Beauty & personalcare") ~ "Chăm sóc sắc đẹp & Chăm sóc cá nhân"
                              , category %in% c("Household Appliances", "Home Suppliers") ~ "Đồ gia dụng"
                              # , category %in% c("Home Suppliers") ~ "Sửa chữa nhà cửa"
                              , category %in% c("Fashion Accessories") ~ "Phụ kiện thời trang"
                              , category %in% c("Shoes", "shoes") ~ "Giày"
                              , category %in% c("Kitchenswear") ~ "Đồ dùng nhà bếp"
                              , category %in% c("Health") ~ "Sức khỏe"
                              , category %in% c("Luggage & Bags","luggage and bags") ~ "Hành lý & Túi xách"
                              , category %in% c("Mom&baby") ~ "Trẻ sơ sinh & thai sản"
                              , category %in% c("Accessories") ~ "Phụ kiện trang sức & Phái sinh"
                              , category %in% c("Phone & Electronic") ~ "Điện thoại & Đồ điện tử"
                              , category %in% c("f&B", "F&B") ~ "Đồ ăn & Đồ uống"
                              , category %in% c("Grocery") ~ "Đồ ăn & Đồ uống"
                              , TRUE ~ "Unknown"))
```

<h5><small>Assumption and Data Sources</h5>

-   Assumption: Revenue is Gross Merchandise Value (GMV).
-   Currency Unit: billion VND (bil VND)
-   Data Source: Data for competitors including TikTok, Shopee, Lazada, Sendo, and Tiki was obtained through Metric.vn's web scraping tool. MeLive's data was provided directly by the MeLive team in a file named "gmv_nmv".</small>

# Executive Summary

<b>MeLive in the Evolving E-commerce Landscape</b>

**Market Overview:**

-   Vietnam's e-commerce market witnessed explosive growth in 2023, surging 78% to 225 trillion VND (9.2 billion USD).
-   Shopee remains the leader but saw a 3% market share decline despite a 68% growth.
-   TikTok Shop experienced a meteoric rise (1355% growth), capturing significant market share (16%) primarily from Lazada and Tiki.
-   Live streaming, popularized by TikTok Shop, is becoming a crucial battleground for e-commerce platforms.
-   Fashion and Health & Beauty are dominant categories on Shopee, contributing a combined 66%.

**MeLive Performance:**

-   In H2 2023, MeLive generated 433 billion VND, representing <b><u>2.1% of TikTok's revenue.</u></b> MeLive's contribution fluctuated but exhibited strong recovery in Q4, achieving double-digit MoM growth against TikTok's decline.
-   MeLive currently holds a roughly <b><u>0.33% share of the total e-commerce market</u></b>.

**Key Category Insights:**

-   Beauty & Personal Care: Dominant on both platforms, but significantly higher (46%) on MeLive compared to TikTok (18.1%).
-   Womenswear: Presents a significant growth opportunity for MeLive, with low contribution (2.4%) despite strong presence on TikTok (20%).
-   Expansion Potential: Categories like Baby, Food & Beverage, and Sport & Outdoors offer promising avenues for MeLive to diversify its portfolio.

**Recommendations:**

-   Leverage MeLive's strong performance in Q4 to maintain momentum and capture a larger share of TikTok's growing e-commerce market.
-   Capitalize on the opportunity in Womenswear by attracting sellers and promoting the category.
-   Explore expansion into high-growth categories like Baby, Food & Beverage, and Sport & Outdoors to diversify MeLive's offerings and address consumer demand.

# Total Revenue

In 2023, total e-commerce sales surged from approximately 127 trillion VND to around 225 trillion VND, marking a remarkable 78% increase. This translates to a growth from 5.2 billion USD to 9.2 billion USD.

Notably, Shopee contributed significantly with a 69% increase, while Tik Tok Shop experienced an extraordinary growth of 1355%.

```{r}
#| echo: false
#| warning: false

total_revenue_2022_2023_tb <- competitor_data_2022_2023 %>% 
  group_by(year, platform) %>% 
  summarise(revenue_bil_vnd = sum(revenue_bil_vnd)) %>% 
  ungroup() %>% 
  spread(year, revenue_bil_vnd) %>% 
  adorn_totals() %>% 
  mutate(`2023_vs_2022` = percent(`2023`/`2022`-1, digits = 1))

knitr::kable(total_revenue_2022_2023_tb
             , digits = 0
             , caption = "VN Ecommerce Revenue And Growth Rate (bil VND)"
             , format.args = list(big.mark = ",", scientific = FALSE)
             , "simple") %>%
  kable_styling(full_width = FALSE, bootstrap_options = "striped")

total_revenue_2022_2023 <- competitor_data_2022_2023 %>% 
  group_by(year, platform) %>% 
  summarise(revenue_bil_vnd = sum(revenue_bil_vnd)) %>% 
  ungroup() %>% 
  mutate(color = case_when(platform == "Shopee" ~ "#ff6d01"
                           , platform == "Tiktok" ~ "#000000"
                           , platform == "Lazada" ~ "#0000ff"
                           , platform == "Tiki" ~ "#4285f4"
                           , platform == "Sendo" ~ "#ea4335"))

CreateGroupedColumnChart(total_revenue_2022_2023
                         , colX = year
                         , colY = revenue_bil_vnd
                         , groupCol = platform
                         , colYTitle = "Revenue (bil VND)") %>%
  hc_colors(total_revenue_2022_2023$color)
```

# Market Share

While Shopee achieved a respectable 68% year-over-year growth, they experienced a 3% decline in market share. In contrast, TikTok Shop captured significant market share, jumping from 2% in 2022 to 16% in 2023. This impressive rise primarily came from Lazada's market share reduction (from 21% to 15%) and Tiki's significant drop (from 4.5% to 1.1%).

Notably, TikTok Shop's livestream-based, entertainment-driven approach has compelled Shopee to invest heavily in live streaming functionality to remain competitive.

```{r}
#| echo: false
#| warning: false

market_share_2022 <- total_revenue_2022_2023 %>% 
  filter(year =="2022") %>% 
  arrange(desc(revenue_bil_vnd)) %>% 
  mutate(color = case_when(platform == "Shopee" ~ "#ff6d01"
                           , platform == "Tiktok" ~ "#000000"
                           , platform == "Lazada" ~ "#0000ff"
                           , platform == "Tiki" ~ "#4285f4"
                           , platform == "Sendo" ~ "#ea4335"))

CreatePieChart(data = market_share_2022
               , colX = platform
               , colY = revenue_bil_vnd) %>% 
  hc_colors(market_share_2022$color) %>% 
  hc_plotOptions(pie = list(
    dataLabels = list(
      enabled = TRUE,
      format = '{point.name}: {point.percentage:.1f} %'))) %>%
  hc_title(text = "VN Market Share 2022")

market_share_2023 <- total_revenue_2022_2023 %>% 
  filter(year =="2023") %>% 
  arrange(desc(revenue_bil_vnd)) %>% 
  mutate(color = case_when(platform == "Shopee" ~ "#ff6d01"
                           , platform == "Tiktok" ~ "#000000"
                           , platform == "Lazada" ~ "#0000ff"
                           , platform == "Tiki" ~ "#4285f4"
                           , platform == "Sendo" ~ "#ea4335"))

CreatePieChart(data = market_share_2023
               , colX = platform
               , colY = revenue_bil_vnd) %>% 
  hc_colors(market_share_2023$color) %>% 
  hc_plotOptions(pie = list(
    dataLabels = list(
      enabled = TRUE,
      format = '{point.name}: {point.percentage:.1f} %'))) %>%
   hc_title(text = "VN Market Share 2023")

```

# Monthly Revenue In 2023

```{r}
#| echo: false
#| warning: false

# Define colors for each category
# category_colors <- list("Shopee" = "#ff6d01"
#                      , "Tiktok" = "#000000"
#                      , "Lazada" = "#0000ff"
#                      , "Tiki" = "#4285f4"
#                      , "Sendo" = "#ea4335"
#                      , "MeLive" = "#dc1a5d")

monthly_revenue <- competitor_data_2023 %>% 
  group_by(month, platform) %>% 
  summarise(revenue_bil_vnd = sum(revenue_bil_vnd)) %>% 
  ungroup()

full_month <- monthly_revenue %>% 
  select(month) %>% 
  distinct()

melive_revenue <- gmv_nmv %>% 
  group_by(month) %>% 
  summarise(revenue_bil_vnd = sum(est_gmv)/10^9) %>% 
  ungroup() %>% 
  right_join(full_month) %>% 
  mutate(platform = "MeLive") %>% 
  mutate_if(is.numeric, replace_na, 0)

monthly_revenue <- monthly_revenue %>% 
  bind_rows(melive_revenue) %>%
  mutate(month = as.Date(month, format = "%Y-%m")) %>% 
  arrange(month, platform) %>% 
  mutate(color = case_when(platform == "Shopee" ~ "#ff6d01"
                           , platform == "Tiktok" ~ "#000000"
                           , platform == "Lazada" ~ "#0000ff"
                           , platform == "Tiki" ~ "#4285f4"
                           , platform == "Sendo" ~ "#ea4335"
                           , platform == "MeLive" ~ "#dc1a5d"))

chart <- monthly_revenue %>%
    hchart(
      'line', hcaes(x = month, y = revenue_bil_vnd, group = platform)
    ) %>%
    hc_yAxis(title = list(text = "Revenue (bil VND)")) %>%
    hc_colors(colors = monthly_revenue$color) %>%
    FormatTooltip(valueDecimal = 0)

chart

```

# Categories

## Shopee

Fashion and Health & Beauty remain the dominant categories on Shopee, comprising 47% of sales.

```{r}
#| echo: false
#| warning: false

shopee_cate <- competitor_shopee_cate %>%
  mutate(percent = revenue_bil_vnd/sum(revenue_bil_vnd)) %>% 
  adorn_totals() %>% 
  mutate(percent = percent(percent, digits = 1))

knitr::kable(shopee_cate
             , digits = 0
             , caption = "Shopee Categories Revenue And Contribution (bil VND)"
             , format.args = list(big.mark = ",", scientific = FALSE)
             , "simple") %>%
  kable_styling(full_width = FALSE, bootstrap_options = "striped")


CreateColumnChart(shopee_cate %>% filter(categories != "Total")
                  , colY = revenue_bil_vnd
                  , colX = categories
                  , colYTitle = "Revenue (bil VND)") %>% 
  hc_colors("#ff6d01")


CreatePieChart(shopee_cate %>% filter(categories != "Total")
                  , colY = revenue_bil_vnd
                  , colX = categories) %>% 
  hc_colors(brewer.pal(length(shopee_cate$categories), 'Spectral')) %>% 
  hc_plotOptions(pie = list(
    dataLabels = list(
      enabled = TRUE,
      format = '{point.name}: {point.percentage:.1f} %'))) %>%
   hc_title(text = "Shopee Categories Contribution 2023")

```

## Tiktok

Contribution of Fashion & Health - Beauty is critical with 66% contribution

```{r}
#| echo: false
#| warning: false

tiktok_cate <- competitor_tiktok_cate %>%
  mutate(percent = revenue_bil_vnd/sum(revenue_bil_vnd)) %>% 
  adorn_totals() %>% 
  mutate(percent = percent(percent, digits = 1))

knitr::kable(tiktok_cate
             , digits = 0
             , caption = "Tiktok Categories Revenue And Contribution (bil VND)"
             , format.args = list(big.mark = ",", scientific = FALSE)
             , "simple") %>%
  kable_styling(full_width = FALSE, bootstrap_options = "striped")


CreateColumnChart(tiktok_cate %>% filter(categories != "Total")
                  , colY = revenue_bil_vnd
                  , colX = categories
                  , colYTitle = "Revenue (bil VND)") %>% 
  hc_colors("#000000")


CreatePieChart(tiktok_cate %>% filter(categories != "Total")
                  , colY = revenue_bil_vnd
                  , colX = categories) %>% 
  hc_colors(brewer.pal(length(tiktok_cate$categories), 'Spectral')) %>% 
  hc_plotOptions(pie = list(
    dataLabels = list(
      enabled = TRUE,
      format = '{point.name}: {point.percentage:.1f} %'))) %>%
   hc_title(text = "Tiktok Categories Contribution 2023")

```

# Melive Vs. Tiktok

## Total revenue in 6 months

-   In the second half of 2023, MeLive generated approximately 433 billion VND, representing 2.1% of TikTok's total revenue. MeLive's contribution then fluctuated, reaching a low of 1.4% in July before recovering to 4.3% by December 2023.

-   During the peak e-commerce season in Q4 2023, MeLive achieved impressive double-digit month-over-month growth, contrasting with TikTok's double-digit decline in revenue share.

::: callout-note
Position Of MeLive in the Ecommerce Market: While MeLive contributes 2.1% to TikTok's e-commerce market share, which holds 15.8% of the total market, MeLive's overall contribution to the total e-commerce market is roughly 0.33%.
:::

```{r}
#| echo: false
#| warning: false


melive_revenue <- gmv_nmv %>% 
  group_by(month) %>% 
  summarise(revenue_bil_vnd = sum(est_gmv)/10^9) %>% 
  ungroup() %>% 
  mutate(platform = "MeLive") %>% 
  mutate_if(is.numeric, replace_na, 0)


tiktok_monthly_revenue <- competitor_data_2023 %>%
  filter(platform == "Tiktok") %>% 
  group_by(month, platform) %>% 
  summarise(revenue_bil_vnd = sum(revenue_bil_vnd)) %>% 
  ungroup() %>% 
  filter(month >= min(melive_revenue$month))

tiktok_monthly_revenue <- tiktok_monthly_revenue %>% 
  bind_rows(melive_revenue) %>%
  mutate(month = as.Date(month, format = "%Y-%m")) %>% 
  arrange(month, desc(revenue_bil_vnd))

tiktok_monthly_revenue_tb <- tiktok_monthly_revenue %>% 
  spread(platform, revenue_bil_vnd) %>% 
  mutate(MeLive_MoM = MeLive/lag(MeLive)-1
         , Tiktok_MoM = Tiktok/lag(Tiktok) -1) %>% 
  select(month, contains('MeLive'), contains('Tiktok')) %>%
  adorn_totals() %>%
  mutate(MeLive_MoM = ifelse(month == "Total", NA_real_, MeLive_MoM)
         , Tiktok_MoM = ifelse(month == "Total", NA_real_, Tiktok_MoM)) %>% 
  mutate(MeLive_vs_Tiktok = percent(MeLive/Tiktok, digits = 1)) %>% 
  mutate_at(vars("MeLive_MoM", "Tiktok_MoM"), percent, digits = 1)

knitr::kable(tiktok_monthly_revenue_tb
             , digits = 0
             , caption = "MeLive Vs Tiktok Revenue And Growth Rate (bil VND)"
             , format.args = list(big.mark = ",", scientific = FALSE)
             , "simple") %>%
  kable_styling(full_width = FALSE, bootstrap_options = "striped")

chart <- tiktok_monthly_revenue %>%
    hchart(
      'line', hcaes(x = month, y = revenue_bil_vnd, group = platform)
    ) %>%
    hc_yAxis(title = list(text = "Revenue (bil VND)")) %>%
    hc_colors(colors = c("#dc1a5d", "#000000")) %>%
    FormatTooltip(valueDecimal = 0)

chart
```

## Contribution By Categories

<i>Data Source Note:

-   TikTok: Revenue by category data comes from Metric.vn. They have cleaned and mapped TikTok categories using Shopee's benchmarks, potentially leading to some inconsistencies.
-   MeLive: Category information comes directly from the MeLive team. This data might be biased due to seller classifications and their primary categories, as sellers can offer products across various categories.

Recommendation:

-   Consider this data a valuable reference, but be aware of potential biases for a more accurate understanding.</i>

Key Takeaways:

-   Beauty & Personal Care: Dominates both platforms, but with a significant difference. It represents 18.1% of TikTok's categories compared to a much higher 46% in MeLive.
-   Womenswear: Presents a growth opportunity for MeLive. While contributing 20% on TikTok, it only accounts for 2.4% of MeLive's categories.
-   Expansion Potential: Categories like Baby, Food & Beverage, and Sport & Outdoors offer promising avenues for MeLive to expand its portfolio.

```{r}
#| echo: false
#| warning: false

summary_category <- gmv_nmv %>% 
  group_by(category) %>% 
  summarise(melive = sum(nmv)/10^9) %>% 
  ungroup() %>% 
  arrange(desc(melive)) %>% 
  full_join(tiktok_rev) %>%
  rename(tiktok = revenue_bil_vnd) %>% 
  mutate_if(is.numeric, replace_na, 0) %>% 
  mutate(percent_ml = percent(melive/sum(melive), digits = 1)) %>% 
  mutate(percent_tiktok = percent(tiktok/sum(tiktok), digits = 1)) %>%
  adorn_totals() %>% 
  mutate(ml_share_in_tiktok = percent(melive/tiktok, digits = 2))

knitr::kable(summary_category
             , digits = 0
             , caption = "GMV By Category vs Tiktok Shop (bil VND)"
             , format.args = list(big.mark = ",", scientific = FALSE)
             , "simple")

CreatePieChart(data = summary_category %>% filter(category != "Total")
               , colX = category
               , colY = melive) %>% 
  hc_title(text = "Melive Revenue Contribution By Category") %>% 
  hc_colors(brewer.pal(length(unique(tiktok_rev$category)), 'Spectral')) %>% 
  hc_plotOptions(pie = list(
    dataLabels = list(
      enabled = TRUE,
      format = '{point.name}: {point.percentage:.1f} %')))

CreatePieChart(data = summary_category %>% filter(category != "Total")
               , colX = category
               , colY = tiktok) %>% 
  hc_title(text = "Tiktok Revenue Contribution By Category") %>% 
  hc_colors(brewer.pal(length(unique(tiktok_rev$category)), 'Spectral')) %>% 
  hc_plotOptions(pie = list(
    dataLabels = list(
      enabled = TRUE,
      format = '{point.name}: {point.percentage:.1f} %')))


```
